Covid-19 A Spatio-Temporal Investigation.

### 2021-01-10 Notebooks/paper_covid.rmd
Update paper_covid.rmd

 The values range from -1 to 1, with the 0 value indicating an absence of spatial autocorrelation, positive values are indicators of positive spatial association, while negative values mean there is an inverse association, since this work seeks to find positive spatial association, the hypothesis tests were built within a 95% confidence level, so that the alternative hypothesis (H1) corresponds to an index higher than zero and the null hypothesis (Ho) to a lack of spatial association, both of which can be written mathematically as Ho=0 and H1>0. The monthly quantities calculated were important to avoid problems of non-stationarity when computing the Global Moran's I, according to Cardoso (2007, p.3, translated from portuguese) "The index loses its validity when calculated for non-stationary data.", in order to validate the index 10,000 simulations were computed in a test of random permutations (*Monte-Carlo*) $\!\!^8$.

@@

 Já o Índice de Moran local (*LISA*) é calculado pela seguinte equação:
 $$I(i) = \frac{x_{i}-\overline x}{s^{2}_{i}}\sum_{j=1}^{n}w_{ij}(x_{j}-\overline{x}), i=1,...,n, j\neq i$$
 Onde $s^{2}_{i}$ é a variância da variável de estudo $x_{i}$:
  $$s^{2}(i) = \frac{\sum_{j=1}^{n}(x_{j}-\overline x)^{2}}{n-1}, i=1,...,n, j\neq i$$
  Os mapas de correlação local de Moran, os *Boxmaps* e as tabelas de moran global, foram feitas para as variáveis (número de recuperados, casos ou mortes, por exemplo) que atingissem correlação global IM>0,50. Os *Boxmaps* são uma representação cartográfica dos quadrantes do gráfico de dispersão de Moran, onde temos que em Seffrin, Araujo e Bazzi (2018,p.28) os valores AA e BB significam respectivamente que essas áreas possuem valores altos rodeadas por áreas que apresentam valores também altos e valores baixos da variável de interesse, rodeadas por áreas que apresentam valores baixos, o valor BA representa áreas com valor baixo rodeada de outras áreas com valores altos, já o AB é o contrário, são regiões com valores altos próximas de locais com valores baixos.

>>>>

 The values range from -1 to 1, with the 0 value indicating an absence of spatial autocorrelation, positive values are indicators of positive spatial association, while negative values mean there is an inverse association, since this work seeks to find positive spatial association, the hypothesis tests were built within a 95% confidence level, so that the alternative hypothesis (H1) corresponds to an index higher than zero and the null hypothesis (Ho) to a lack of spatial association, both of which can be written mathematically as Ho=0 and H1>0. The monthly quantities were important to avoid problems of non-stationarity when computing the Global Moran's I, according to Cardoso (2007, p.3, translated from portuguese) "The index loses its validity when calculated for non-stationary data.", in order to validate the index 10,000 simulations were computed in a test of random permutations (*Monte-Carlo*) $\!\!^8$.

@@

**Equation 2**
 $$I(i) = \frac{x_{i}-\overline x}{s^{2}_{i}}\sum_{j=1}^{n}w_{ij}(x_{j}-\overline{x}), i=1,...,n, j\neq i$$
 Onde $s^{2}_{i}$ é a variância da variável de estudo $x_{i}$:
  $$s^{2}(i) = \frac{\sum_{j=1}^{n}(x_{j}-\overline x)^{2}}{n-1}, i=1,...,n, j\neq i$$
  Os mapas de correlação local de Moran, os *Boxmaps* e as tabelas de moran global, foram feitas para as variáveis (número de recuperados, casos ou mortes, por exemplo) que atingissem correlação global IM>0,50. Os *Boxmaps* são uma representação cartográfica dos quadrantes do gráfico de dispersão de Moran, onde temos que em Seffrin, Araujo e Bazzi (2018,p.28) os valores AA e BB significam respectivamente que essas áreas possuem valores altos rodeadas por áreas que apresentam valores também altos e valores baixos da variável de interesse, rodeadas por áreas que apresentam valores baixos, o valor BA representa áreas com valor baixo rodeada de outras áreas com valores altos, já o AB é o contrário, são regiões com valores altos próximas de locais com valores baixos.
